---
title: "Property Listing Data Analysis"
author: "alc00"
date: "12/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Connecting to the Database

## Loading the Necessary Libraries
```{r, message=FALSE}
# dplyr - Data Manipulation and Core Module for dbplyr
library(dplyr)

# dbplyr - Required for Connection to SQLite Database
library(dbplyr)

# RSQLite - Required for Connection to SQLite Database
library(RSQLite)

# sqldf - Allows you to use SQL for R objects
library(sqldf)

# Caret - Useful Package for Predictive Modeling
library(caret)

# rjson - Reads JSON data
library(rjson)
```

## Connecting to SQLite Database

```{r}
db_file_path <- "C:/Users/Allen/Desktop/Code/Projects/Property/SQLite/project_property.db"

db <- DBI::dbConnect(RSQLite::SQLite(), db_file_path)
```

## Check Database Connection
```{r}
src_dbi(db)
```
# Data Analysis

## Query and Clean Data using SQL and dump data as CSV for Tableau Public
```{r}
DM_99_CO_LISTING_SUMMARY_T <- tbl(db, sql(
                      "
                      SELECT
                      DTIME_INSERTED,
                      LISTING_TYPE,
                      RENTAL_TYPE,
                      LISTING_ID,
                      LISTING_NAME,
                      VALUE,
                      VALUE_PSF,
                      CASE WHEN CNT_BEDS = 'N/A' THEN NULL ELSE CNT_BEDS END AS CNT_BEDS,
                      CASE WHEN CNT_BATHS = 'N/A' THEN NULL ELSE CNT_BATHS END AS CNT_BATHS,
                      SIZE_SQFT,
                      LATITUDE,
                      LONGITUDE,
                      LISTING_LINK
                      FROM DM_99_CO_LISTING_SUMMARY_T
                      "))

DM_99_CO_LISTING_ADD_INFO_T <- tbl(db, sql(
                      "
                      SELECT 
                      * 
                      FROM DM_99_CO_LISTING_ADD_INFO_T
                      "))

DM_99_CO_LISTING_PROPERTY_T <- tbl(db, sql(
                      "
                      SELECT 
                      * 
                      FROM DM_99_CO_LISTING_PROPERTY_T
                      "))

DM_99_CO_LISTING_AMENITIES_T <- tbl(db, sql(
                      "
                      SELECT * FROM DM_99_CO_LISTING_AMENITIES_T
                      "))

write.csv(DM_99_CO_LISTING_SUMMARY_T, "C:/Users/Allen/Desktop/Code/Projects/Property/Tableau Public Data/DM_99_CO_LISTING_SUMMARY_T.csv")

write.csv(DM_99_CO_LISTING_ADD_INFO_T, "C:/Users/Allen/Desktop/Code/Projects/Property/Tableau Public Data/DM_99_CO_LISTING_ADD_INFO_T.csv")

write.csv(DM_99_CO_LISTING_PROPERTY_T, "C:/Users/Allen/Desktop/Code/Projects/Property/Tableau Public Data/DM_99_CO_LISTING_PROPERTY_T.csv")

write.csv(DM_99_CO_LISTING_AMENITIES_T, "C:/Users/Allen/Desktop/Code/Projects/Property/Tableau Public Data/DM_99_CO_LISTING_AMENITIES_T.csv")
```

## Temporary Fix: SQFT Fix
```{r}
DM_99_CO_LISTING_SUMMARY_T <- DM_99_CO_LISTING_SUMMARY_T %>% mutate(SQFT_FIX = VALUE/VALUE_PSF)
```

## Addition of Descriptive Statistics for Use in Predictive Model

### Remaining Government Lease
```{r}
current_year_int <- as.numeric(as.character(Sys.Date(),"%Y"))

DM_99_CO_LISTING_PROPERTY_T <-
DM_99_CO_LISTING_PROPERTY_T %>% mutate(
    remaining_gov_lease =
      if(GOV_LEASE == "N/A" | BUILD_YEAR == "N/A" | GOV_LEASE == "-") {
      NULL
    } else if (GOV_LEASE == "Freehold") {
      "Freehold"
    } else {
      as.character((as.numeric(BUILD_YEAR) + as.numeric(GOV_LEASE)) - current_year_int)
    })
```

### Amenities Count and Indicators
```{r}

# Check Top Amenities for Sale Listings
tbl(db, sql(
            "
            SELECT 
            A.NAME_AMENITY,
            COUNT(1) CNT
            FROM DM_99_CO_LISTING_AMENITIES_T A
            JOIN DM_99_CO_LISTING_SUMMARY_T B
              ON A.LISTING_ID = B.LISTING_ID
            WHERE B.LISTING_TYPE = 'Sale'
            GROUP BY
            A.NAME_AMENITY
            ORDER BY 2 DESC
            "))

# Check Top Amenities for Rental Unit Listings
tbl(db, sql(
            "
            SELECT 
            A.NAME_AMENITY,
            COUNT(1) CNT
            FROM DM_99_CO_LISTING_AMENITIES_T A
            JOIN DM_99_CO_LISTING_SUMMARY_T B
              ON A.LISTING_ID = B.LISTING_ID
            WHERE B.LISTING_TYPE = 'Rental'
            AND B.RENTAL_TYPE = 'Entire Unit/Property'
            GROUP BY
            A.NAME_AMENITY
            ORDER BY 2 DESC
            "))

# Check Top Amenities for Rental Room Listings
tbl(db, sql(
            "
            SELECT 
            A.NAME_AMENITY,
            COUNT(1) CNT
            FROM DM_99_CO_LISTING_AMENITIES_T A
            JOIN DM_99_CO_LISTING_SUMMARY_T B
              ON A.LISTING_ID = B.LISTING_ID
            WHERE B.LISTING_TYPE = 'Rental'
            AND B.RENTAL_TYPE <> 'Entire Unit/Property'
            GROUP BY
            A.NAME_AMENITY
            ORDER BY 2 DESC
            "))


# Get Relevant Amenities for Sale Listings
listing_sale_amenities <-
  tbl(db, sql(
            "
            SELECT 
            A.LISTING_ID,
            COUNT(DISTINCT A.NAME_AMENITY) CNT_AMENITIES,
            SUM(CASE WHEN A.NAME_AMENITY = 'Gym' THEN 1 ELSE 0 END) FLAG_GYM,
            SUM(CASE WHEN A.NAME_AMENITY = 'Balcony' THEN 1 ELSE 0 END) FLAG_BALCONY,
            SUM(CASE WHEN A.NAME_AMENITY = 'Playground' THEN 1 ELSE 0 END) FLAG_PLAYGROUND,
            SUM(CASE WHEN A.NAME_AMENITY = 'Security' THEN 1 ELSE 0 END) FLAG_SECURITY,
            SUM(CASE WHEN A.NAME_AMENITY = 'Swimming Pool' THEN 1 ELSE 0 END) FLAG_POOL,
            SUM(CASE WHEN A.NAME_AMENITY = 'Renovated' THEN 1 ELSE 0 END) FLAG_RENOVATED,
            SUM(CASE WHEN A.NAME_AMENITY = 'Parking' THEN 1 ELSE 0 END) FLAG_PARKING,
            SUM(CASE WHEN A.NAME_AMENITY = 'Jacuzzi' THEN 1 ELSE 0 END) FLAG_JACUZZI,
            SUM(CASE WHEN A.NAME_AMENITY = 'Fibre Ready' THEN 1 ELSE 0 END) FLAG_FIBRE,
            SUM(CASE WHEN A.NAME_AMENITY = 'City View' THEN 1 ELSE 0 END) FLAG_CITY_VIEW,
            SUM(CASE WHEN A.NAME_AMENITY = 'Pool View' THEN 1 ELSE 0 END) FLAG_POOL_VIEW,
            SUM(CASE WHEN A.NAME_AMENITY = 'Maid Room' THEN 1 ELSE 0 END) FLAG_MAID_ROOM,
            SUM(CASE WHEN A.NAME_AMENITY = 'Study Room' THEN 1 ELSE 0 END) FLAG_STUDY_ROOM,
            SUM(CASE WHEN A.NAME_AMENITY = 'High Floor' THEN 1 ELSE 0 END) FLAG_HIGH_FLOOR,
            SUM(CASE WHEN A.NAME_AMENITY = 'Mid Floor' THEN 1 ELSE 0 END) FLAG_MID_FLOOR,
            SUM(CASE WHEN A.NAME_AMENITY = 'Low Floor' THEN 1 ELSE 0 END) FLAG_LOW_FLOOR,
            SUM(CASE WHEN A.NAME_AMENITY = 'Top Floor' THEN 1 ELSE 0 END) FLAG_TOP_FLOOR,
            SUM(CASE WHEN A.NAME_AMENITY = 'Penthouse' THEN 1 ELSE 0 END) FLAG_PENTHOUSE
            FROM DM_99_CO_LISTING_AMENITIES_T A
            JOIN DM_99_CO_LISTING_SUMMARY_T B
              ON A.LISTING_ID = B.LISTING_ID
            WHERE B.LISTING_TYPE = 'Sale'
            GROUP BY A.LISTING_ID
            "))

# Get Relevant Amenities for Rental Unit Listings
listing_rental_unit_amenities <-
  tbl(db, sql(
            "
            SELECT 
            A.LISTING_ID,
            COUNT(DISTINCT A.NAME_AMENITY) CNT_AMENITIES,
            SUM(CASE WHEN A.NAME_AMENITY = 'Aircon' THEN 1 ELSE 0 END) FLAG_AIRCON,
            SUM(CASE WHEN A.NAME_AMENITY = 'Fridge' THEN 1 ELSE 0 END) FLAG_FRIDGE,
            SUM(CASE WHEN A.NAME_AMENITY = 'Washer' THEN 1 ELSE 0 END) FLAG_WASHER,
            SUM(CASE WHEN A.NAME_AMENITY = 'Stove' THEN 1 ELSE 0 END) FLAG_STOVE,
            SUM(CASE WHEN A.NAME_AMENITY = 'Oven' THEN 1 ELSE 0 END) FLAG_OVEN,
            SUM(CASE WHEN A.NAME_AMENITY = 'Dryer' THEN 1 ELSE 0 END) FLAG_DRYER,
            SUM(CASE WHEN A.NAME_AMENITY = 'Bed' THEN 1 ELSE 0 END) FLAG_BED,
            SUM(CASE WHEN A.NAME_AMENITY = 'Closet' THEN 1 ELSE 0 END) FLAG_CLOSET,
            SUM(CASE WHEN A.NAME_AMENITY = 'Sofa' THEN 1 ELSE 0 END) FLAG_SOFA,
            SUM(CASE WHEN A.NAME_AMENITY = 'Dining Room Furniture	' THEN 1 ELSE 0 END) FLAG_DINING,
            SUM(CASE WHEN A.NAME_AMENITY = 'Bathtub' THEN 1 ELSE 0 END) FLAG_BATHTUB,
            SUM(CASE WHEN A.NAME_AMENITY = 'TV' THEN 1 ELSE 0 END) FLAG_TV,
            SUM(CASE WHEN A.NAME_AMENITY = 'Gym' THEN 1 ELSE 0 END) FLAG_GYM,
            SUM(CASE WHEN A.NAME_AMENITY = 'Balcony' THEN 1 ELSE 0 END) FLAG_BALCONY,
            SUM(CASE WHEN A.NAME_AMENITY = 'Playground' THEN 1 ELSE 0 END) FLAG_PLAYGROUND,
            SUM(CASE WHEN A.NAME_AMENITY = 'Security' THEN 1 ELSE 0 END) FLAG_SECURITY,
            SUM(CASE WHEN A.NAME_AMENITY = 'Swimming Pool' THEN 1 ELSE 0 END) FLAG_POOL,
            SUM(CASE WHEN A.NAME_AMENITY = 'Renovated' THEN 1 ELSE 0 END) FLAG_RENOVATED,
            SUM(CASE WHEN A.NAME_AMENITY = 'Parking' THEN 1 ELSE 0 END) FLAG_PARKING,
            SUM(CASE WHEN A.NAME_AMENITY = 'Jacuzzi' THEN 1 ELSE 0 END) FLAG_JACUZZI,
            SUM(CASE WHEN A.NAME_AMENITY = 'Fibre Ready' THEN 1 ELSE 0 END) FLAG_FIBRE,
            SUM(CASE WHEN A.NAME_AMENITY = 'City View' THEN 1 ELSE 0 END) FLAG_CITY_VIEW,
            SUM(CASE WHEN A.NAME_AMENITY = 'Pool View' THEN 1 ELSE 0 END) FLAG_POOL_VIEW,
            SUM(CASE WHEN A.NAME_AMENITY = 'Maid Room' THEN 1 ELSE 0 END) FLAG_MAID_ROOM,
            SUM(CASE WHEN A.NAME_AMENITY = 'Study Room' THEN 1 ELSE 0 END) FLAG_STUDY_ROOM,
            SUM(CASE WHEN A.NAME_AMENITY = 'High Floor' THEN 1 ELSE 0 END) FLAG_HIGH_FLOOR,
            SUM(CASE WHEN A.NAME_AMENITY = 'Mid Floor' THEN 1 ELSE 0 END) FLAG_MID_FLOOR,
            SUM(CASE WHEN A.NAME_AMENITY = 'Low Floor' THEN 1 ELSE 0 END) FLAG_LOW_FLOOR,
            SUM(CASE WHEN A.NAME_AMENITY = 'Top Floor' THEN 1 ELSE 0 END) FLAG_TOP_FLOOR,
            SUM(CASE WHEN A.NAME_AMENITY = 'Penthouse' THEN 1 ELSE 0 END) FLAG_PENTHOUSE
            FROM DM_99_CO_LISTING_AMENITIES_T A
            JOIN DM_99_CO_LISTING_SUMMARY_T B
              ON A.LISTING_ID = B.LISTING_ID
            WHERE B.LISTING_TYPE = 'Rental'
            AND B.RENTAL_TYPE = 'Entire Unit/Property'
            GROUP BY A.LISTING_ID
            "))


# Get Relevant Amenities for Rental Room Listings
listing_rental_room_amenities <- 
  tbl(db, sql(
            "
            SELECT 
            A.LISTING_ID,
            COUNT(DISTINCT A.NAME_AMENITY) CNT_AMENITIES,
            SUM(CASE WHEN A.NAME_AMENITY = 'Aircon' THEN 1 ELSE 0 END) FLAG_AIRCON,
            SUM(CASE WHEN A.NAME_AMENITY = 'Fridge' THEN 1 ELSE 0 END) FLAG_FRIDGE,
            SUM(CASE WHEN A.NAME_AMENITY = 'Washer' THEN 1 ELSE 0 END) FLAG_WASHER,
            SUM(CASE WHEN A.NAME_AMENITY = 'Stove' THEN 1 ELSE 0 END) FLAG_STOVE,
            SUM(CASE WHEN A.NAME_AMENITY = 'Oven' THEN 1 ELSE 0 END) FLAG_OVEN,
            SUM(CASE WHEN A.NAME_AMENITY = 'Dryer' THEN 1 ELSE 0 END) FLAG_DRYER,
            SUM(CASE WHEN A.NAME_AMENITY = 'Bed' THEN 1 ELSE 0 END) FLAG_BED,
            SUM(CASE WHEN A.NAME_AMENITY = 'Closet' THEN 1 ELSE 0 END) FLAG_CLOSET,
            SUM(CASE WHEN A.NAME_AMENITY = 'Sofa' THEN 1 ELSE 0 END) FLAG_SOFA,
            SUM(CASE WHEN A.NAME_AMENITY = 'Dining Room Furniture	' THEN 1 ELSE 0 END) FLAG_DINING,
            SUM(CASE WHEN A.NAME_AMENITY = 'Bathtub' THEN 1 ELSE 0 END) FLAG_BATHTUB,
            SUM(CASE WHEN A.NAME_AMENITY = 'TV' THEN 1 ELSE 0 END) FLAG_TV,
            SUM(CASE WHEN A.NAME_AMENITY = 'Wireless Internet' THEN 1 ELSE 0 END) FLAG_WIFI,
            SUM(CASE WHEN A.NAME_AMENITY = 'Attached Bathroom' THEN 1 ELSE 0 END) FLAG_BATHROOM,
            SUM(CASE WHEN A.NAME_AMENITY = 'Gym' THEN 1 ELSE 0 END) FLAG_GYM,
            SUM(CASE WHEN A.NAME_AMENITY = 'Balcony' THEN 1 ELSE 0 END) FLAG_BALCONY,
            SUM(CASE WHEN A.NAME_AMENITY = 'Playground' THEN 1 ELSE 0 END) FLAG_PLAYGROUND,
            SUM(CASE WHEN A.NAME_AMENITY = 'Security' THEN 1 ELSE 0 END) FLAG_SECURITY,
            SUM(CASE WHEN A.NAME_AMENITY = 'Swimming Pool' THEN 1 ELSE 0 END) FLAG_POOL,
            SUM(CASE WHEN A.NAME_AMENITY = 'Renovated' THEN 1 ELSE 0 END) FLAG_RENOVATED,
            SUM(CASE WHEN A.NAME_AMENITY = 'Parking' THEN 1 ELSE 0 END) FLAG_PARKING,
            SUM(CASE WHEN A.NAME_AMENITY = 'Jacuzzi' THEN 1 ELSE 0 END) FLAG_JACUZZI,
            SUM(CASE WHEN A.NAME_AMENITY = 'Fibre Ready' THEN 1 ELSE 0 END) FLAG_FIBRE,
            SUM(CASE WHEN A.NAME_AMENITY = 'City View' THEN 1 ELSE 0 END) FLAG_CITY_VIEW,
            SUM(CASE WHEN A.NAME_AMENITY = 'Pool View' THEN 1 ELSE 0 END) FLAG_POOL_VIEW,
            SUM(CASE WHEN A.NAME_AMENITY = 'Maid Room' THEN 1 ELSE 0 END) FLAG_MAID_ROOM,
            SUM(CASE WHEN A.NAME_AMENITY = 'Study Room' THEN 1 ELSE 0 END) FLAG_STUDY_ROOM,
            SUM(CASE WHEN A.NAME_AMENITY = 'High Floor' THEN 1 ELSE 0 END) FLAG_HIGH_FLOOR,
            SUM(CASE WHEN A.NAME_AMENITY = 'Mid Floor' THEN 1 ELSE 0 END) FLAG_MID_FLOOR,
            SUM(CASE WHEN A.NAME_AMENITY = 'Low Floor' THEN 1 ELSE 0 END) FLAG_LOW_FLOOR,
            SUM(CASE WHEN A.NAME_AMENITY = 'Top Floor' THEN 1 ELSE 0 END) FLAG_TOP_FLOOR,
            SUM(CASE WHEN A.NAME_AMENITY = 'Penthouse' THEN 1 ELSE 0 END) FLAG_PENTHOUSE
            FROM DM_99_CO_LISTING_AMENITIES_T A
            JOIN DM_99_CO_LISTING_SUMMARY_T B
              ON A.LISTING_ID = B.LISTING_ID
            WHERE B.LISTING_TYPE = 'Rental'
            AND B.RENTAL_TYPE <> 'Entire Unit/Property'
            GROUP BY A.LISTING_ID
            "))

head(listing_rental_room_amenities)
head(listing_rental_unit_amenities)
head(listing_sale_amenities)

```
### Convert SQLite Connection Tables to Dataframe and Basic Data Type Changes
```{r}
listing_summary_df <- data.frame(DM_99_CO_LISTING_SUMMARY_T)
listing_add_info_df <- data.frame(DM_99_CO_LISTING_ADD_INFO_T)
listing_property_df <- data.frame(DM_99_CO_LISTING_PROPERTY_T)

listing_rental_room_amenities_df <- data.frame(listing_rental_room_amenities)
listing_rental_unit_amenities_df <- data.frame(listing_rental_unit_amenities)
listing_sale_amenities_df <- data.frame(listing_sale_amenities)


listing_summary_df[,"CNT_BEDS"] = as.character(listing_summary_df[,"CNT_BEDS"])
listing_summary_df[,"CNT_BATHS"] = as.character(listing_summary_df[,"CNT_BATHS"])

```


### Add Location Data based on GPS Data using Photon Geocoding API
```{r}

countrycode <- vector()
county <- vector()
district <- vector()
postcode <- vector()

row_count <- nrow(listing_summary_df)

for (i in 1:row_count) {
  long <- (listing_summary_df[i, "LONGITUDE"])
  lat <- (listing_summary_df[i, "LATITUDE"])
  
  url_rev_geocode <- paste("https://photon.komoot.io/reverse?lon=", long, "&lat=", lat, sep="")

  raw_json_rev_geocode <- fromJSON(file = url_rev_geocode)
  
  countrycode <- append(countrycode, 
                        ifelse(is.null(raw_json_rev_geocode$features[[1]]$properties$countrycode),
                               "N/A",
                               raw_json_rev_geocode$features[[1]]$properties$countrycode
                        ))
  
  county <- append(county,
                        ifelse(is.null(raw_json_rev_geocode$features[[1]]$properties$county),
                               "N/A",
                               raw_json_rev_geocode$features[[1]]$properties$county
                        ))
  
  district <- append(district, 
                        ifelse(is.null(raw_json_rev_geocode$features[[1]]$properties$district),
                               "N/A",
                               raw_json_rev_geocode$features[[1]]$properties$district
                        ))
  
  postcode <- append(postcode, 
                        ifelse(is.null(raw_json_rev_geocode$features[[1]]$properties$postcode),
                               "N/A",
                               raw_json_rev_geocode$features[[1]]$properties$postcode
                        ))
  
  print(paste("row", i, "of", row_count))
}

listing_summary_df <- cbind(listing_summary_df, data.frame(countrycode = countrycode))
listing_summary_df <- cbind(listing_summary_df, data.frame(county = county))
listing_summary_df <- cbind(listing_summary_df, data.frame(district = district))
listing_summary_df <- cbind(listing_summary_df, data.frame(postcode = postcode))

```


## Create Predictive Model to get Predicted Listing Value and Value PSF
The goal would be to create a model that would provide the market listing price and price per square foot based on other listings with similar attributes. This would help determine if the listing is over or under market value.

Note that this will be based on listing price rather than actual price (price sold to tenant/buyer), due to data limitations. This will assume that listing price is near the actual price or that the difference between the actual price and the listing price is relatively consistent.

Predictions:
I. Sales
  1) Estimated Property Valuation (Value)
  2) Estimated Property Valuation per Square Foot (Value PSF)
II. Rental
  1) Estimated Rental Valuation (Value)
  2) Estimated Rental Valuation per Square Foot (Value PSF)
  3) Estimated Property Valuation (Value - Property)
  4) Estimated Property Valuation per Square Foot (Value PSF - Property)

### Sale Listings

#### Join and Clean Tables
```{r}
sale_predictive_data_raw <- 
  sqldf("SELECT
        A.LISTING_ID,
        A.VALUE,
        A.VALUE_PSF,
        A.SQFT_FIX SIZE_SQFT,
        C.TYPE BUILDING_TYPE,
        A.DISTRICT DISTRICT,
        CASE WHEN C.remaining_gov_lease IS NULL THEN 'NO DATA'
             WHEN C.remaining_gov_lease = 'Freehold' THEN 'Freehold'
             WHEN C.remaining_gov_lease <= 20 THEN '0-20 YEARS LEFT'
             WHEN C.remaining_gov_lease <= 40 THEN '21-40 YEARS LEFT'
             WHEN C.remaining_gov_lease <= 60 THEN '41-60 YEARS LEFT'
             WHEN C.remaining_gov_lease <= 80 THEN '61-80 YEARS LEFT'
             WHEN C.remaining_gov_lease <= 99 THEN '81-99 YEARS LEFT'
             ELSE 'MORE THAN 99 YEARS LEFT'
             END AS GOV_LEASE_DURATION
        FROM listing_summary_df A
        LEFT JOIN listing_add_info_df B
          ON A.LISTING_ID = B.LISTING_ID
        LEFT JOIN listing_property_df C
          ON B.PROPERTY_ID = C.PROPERTY_ID
        LEFT JOIN listing_sale_amenities_df D
          ON A.LISTING_ID = D.LISTING_ID
        WHERE A.LISTING_TYPE = 'Sale'
        ")

sale_predictive_data_raw[,grepl('FLAG', colnames(sale_predictive_data_raw))] <-
  sale_predictive_data_raw[,grepl('FLAG', colnames(sale_predictive_data_raw))] %>%
    lapply(FUN = as.character)

summary(sale_predictive_data_raw)

```

#### Create Random Forest Model to fit variables to Price per Square Foot
```{r}
set.seed(20201217)

sales_model_variables <- 
  sale_predictive_data_raw[, !colnames(sale_predictive_data_raw) %in% c("LISTING_ID", "VALUE", "SIZE_SQFT")]

sales_model_rf_c <- train(VALUE_PSF~., data=sales_model_variables, method = "rf")

model_results <- data.frame(cbind(sale_predictive_data_raw$VALUE_PSF, predict(sales_model_rf_c,sales_model_variables)))

colnames(model_results) <- c("actual", "fitted")

model_results[,"Difference"] <- model_results$fitted - model_results$actual

diff_results <- mean(abs(model_results$Difference/model_results$actual))
```

First Iteration: Average deviation of fitted values to actual values = 16.09%

#### Attach Results to Data - Sales
```{r}
listing_id_sale_val_results <- cbind(sale_predictive_data_raw$LISTING_ID, model_results[,2],                                    ifelse(abs(model_results$Difference/model_results$actual) <=
                                            diff_results,"Within Range", "Outside Range"))

colnames(listing_id_sale_val_results) <- c("LISTING_ID", "FITTED_PROPERTY_PSF", "PSF_RF_RANGE")

listing_id_sale_val_results <- data.frame(listing_id_sale_val_results)

```

### Rental Unit Listings

#### Join and Clean Data
```{r}
rental_unit_predictive_data_raw <- 
  sqldf("SELECT
        A.LISTING_ID,
        A.VALUE,
        A.VALUE_PSF,
        A.SQFT_FIX SIZE_SQFT,
        C.TYPE BUILDING_TYPE,
        A.DISTRICT DISTRICT,
        CASE WHEN C.remaining_gov_lease IS NULL THEN 'NO DATA'
           WHEN C.remaining_gov_lease = 'Freehold' THEN 'Freehold'
           WHEN C.remaining_gov_lease <= 20 THEN '0-20 YEARS LEFT'
           WHEN C.remaining_gov_lease <= 40 THEN '21-40 YEARS LEFT'
           WHEN C.remaining_gov_lease <= 60 THEN '41-60 YEARS LEFT'
           WHEN C.remaining_gov_lease <= 80 THEN '61-80 YEARS LEFT'
           WHEN C.remaining_gov_lease <= 99 THEN '81-99 YEARS LEFT'
           ELSE 'MORE THAN 99 YEARS LEFT'
           END AS GOV_LEASE_DURATION
        FROM listing_summary_df A
        LEFT JOIN listing_add_info_df B
          ON A.LISTING_ID = B.LISTING_ID
        LEFT JOIN listing_property_df C
          ON B.PROPERTY_ID = C.PROPERTY_ID
        LEFT JOIN listing_rental_unit_amenities_df D
          ON A.LISTING_ID = D.LISTING_ID
        WHERE A.LISTING_TYPE = 'Rental'
        AND A.RENTAL_TYPE = 'Entire Unit/Property'
        ")

rental_unit_predictive_data_raw[,grepl('FLAG', colnames(rental_unit_predictive_data_raw))] <-
  rental_unit_predictive_data_raw[,grepl('FLAG', colnames(rental_unit_predictive_data_raw))] %>%
    lapply(FUN = as.character)

summary(rental_unit_predictive_data_raw)
```

#### Create Random Forest Model to fit variables to Price per Square Foot
```{r}
set.seed(20201217)

rent_unit_model_variables <- 
  rental_unit_predictive_data_raw[, !colnames(rental_unit_predictive_data_raw) %in% c("LISTING_ID", "VALUE", "SIZE_SQFT")]

rental_unit_model_rf_c <- train(VALUE_PSF~., data=rent_unit_model_variables, method = "rf")

model_results_1 <- data.frame(cbind(rental_unit_predictive_data_raw$VALUE_PSF, predict(rental_unit_model_rf_c,rent_unit_model_variables)))

colnames(model_results_1) <- c("actual", "fitted")

model_results_1[,"Difference"] <- model_results_1$fitted - model_results_1$actual

diff_results <- mean(abs(model_results_1$Difference/model_results_1$actual))
```

Initial Iteration: 24.47% average deviation from fitted values

#### Create Random Forest Model to Estimate Property Valuation
```{r}
model_results_2 <- data.frame(cbind(rental_unit_predictive_data_raw$VALUE_PSF, predict(sales_model_rf_c,rent_unit_model_variables)))
```

#### Attach Results to Data - Rental Units
```{r}
listing_id_ru_val_results <- cbind(rental_unit_predictive_data_raw$LISTING_ID, model_results_2[,2], 
                                   ifelse(abs(model_results_1$Difference/model_results_1$actual) <=
                                            diff_results,"Within Range", "Outside Range"))

colnames(listing_id_ru_val_results) <- c("LISTING_ID", "FITTED_PROPERTY_PSF", "PSF_RF_RANGE")

listing_id_ru_val_results <- data.frame(listing_id_ru_val_results)

```

### Rental Room Listings

#### Join and Clean Data
```{r}
rental_room_predictive_data_raw <- 
  sqldf("SELECT
        A.LISTING_ID,
        A.VALUE,
        A.VALUE_PSF,
        A.SQFT_FIX SIZE_SQFT,
        C.TYPE BUILDING_TYPE,
        A.DISTRICT DISTRICT,
        CASE WHEN C.remaining_gov_lease IS NULL THEN 'NO DATA'
             WHEN C.remaining_gov_lease = 'Freehold' THEN 'Freehold'
             WHEN C.remaining_gov_lease <= 20 THEN '0-20 YEARS LEFT'
             WHEN C.remaining_gov_lease <= 40 THEN '21-40 YEARS LEFT'
             WHEN C.remaining_gov_lease <= 60 THEN '41-60 YEARS LEFT'
             WHEN C.remaining_gov_lease <= 80 THEN '61-80 YEARS LEFT'
             WHEN C.remaining_gov_lease <= 99 THEN '81-99 YEARS LEFT'
             ELSE 'MORE THAN 99 YEARS LEFT'
             END AS GOV_LEASE_DURATION
        FROM listing_summary_df A
        LEFT JOIN listing_add_info_df B
          ON A.LISTING_ID = B.LISTING_ID
        LEFT JOIN listing_property_df C
          ON B.PROPERTY_ID = C.PROPERTY_ID
        LEFT JOIN listing_rental_room_amenities_df D
          ON A.LISTING_ID = D.LISTING_ID
        WHERE A.LISTING_TYPE = 'Rental'
        AND A.RENTAL_TYPE <> 'Entire Unit/Property'
        ")

rental_room_predictive_data_raw[,grepl('FLAG', colnames(rental_room_predictive_data_raw))] <-
  rental_room_predictive_data_raw[,grepl('FLAG', colnames(rental_room_predictive_data_raw))] %>%
    lapply(FUN = as.character)

summary(rental_room_predictive_data_raw)
```

#### Create Random Forest Model to fit variables to Price per Square Foot
```{r}
set.seed(20201217)

rent_room_model_variables <- 
  rental_room_predictive_data_raw[, !colnames(rental_room_predictive_data_raw) %in% c("LISTING_ID", "VALUE", "SIZE_SQFT")]

rental_room_model_rf_c <- train(VALUE_PSF~., data=rent_room_model_variables, method = "rf")

model_results_1 <- data.frame(cbind(rent_room_model_variables$VALUE_PSF, predict(rental_room_model_rf_c,rent_room_model_variables)))

colnames(model_results_1) <- c("actual", "fitted")

model_results_1[,"Difference"] <- model_results_1$fitted - model_results_1$actual

diff_results <- mean(abs(model_results_1$Difference/ifelse(model_results_1$actual==0,1,model_results_1$actual)))
```

Initial Iteration: 82.62% average deviation from fitted values


#### Create Random Forest Model to Estimate Property Valuation
```{r}

# Replace Districts with More Generalized Area due to limitations in Model
rent_room_model_variables[rent_room_model_variables$DISTRICT=="Kallang Way", "DISTRICT"] = "Kallang"
rent_room_model_variables[rent_room_model_variables$DISTRICT=="Serangoon North", "DISTRICT"] = "Serangoon"

model_results_2 <- data.frame(cbind(rent_room_model_variables$VALUE_PSF, predict(sales_model_rf_c,rent_room_model_variables)))


```


#### Attach Results to Data - Rental Rooms
```{r}
listing_id_rr_val_results <- cbind(rental_room_predictive_data_raw$LISTING_ID, model_results_2[,2], 
                                   ifelse(abs(model_results_1$Difference/model_results_1$actual) <=
                                            diff_results,"Within Range", "Outside Range"))

colnames(listing_id_rr_val_results) <- c("LISTING_ID", "FITTED_PROPERTY_PSF", "PSF_RF_RANGE")

listing_id_rr_val_results <- data.frame(listing_id_rr_val_results)

```


### Consolidate Results
```{r}
listing_summary_df_2 <- sqldf(
  "
  SELECT
  A.*,
  CASE WHEN B.FITTED_PROPERTY_PSF IS NOT NULL THEN B.FITTED_PROPERTY_PSF
       WHEN C.FITTED_PROPERTY_PSF IS NOT NULL THEN C.FITTED_PROPERTY_PSF
       WHEN D.FITTED_PROPERTY_PSF IS NOT NULL THEN D.FITTED_PROPERTY_PSF
       ELSE NULL
       END AS FITTED_PROPERTY_PSF,
  CASE WHEN B.PSF_RF_RANGE IS NOT NULL THEN B.PSF_RF_RANGE
       WHEN C.PSF_RF_RANGE IS NOT NULL THEN C.PSF_RF_RANGE
       WHEN D.PSF_RF_RANGE IS NOT NULL THEN D.PSF_RF_RANGE
       ELSE NULL
       END AS PSF_RF_RANGE
  FROM listing_summary_df A
  LEFT JOIN listing_id_sale_val_results B
    ON A.LISTING_ID = B.LISTING_ID
  LEFT JOIN listing_id_ru_val_results C
    ON A.LISTING_ID = C.LISTING_ID
  LEFT JOIN listing_id_rr_val_results D
    ON A.LISTING_ID = D.LISTING_ID
  "
)

write.csv(listing_summary_df_2, "C:/Users/Allen/Desktop/Code/Projects/Property/Tableau Public Data/DM_99_CO_LISTING_SUMMARY_T.csv")
```

